---
title: "Diet Synthesis C Model Comparisons"
output: html_document
---

## About

Here we undertake model comparisons for traitglms.

## Workspace

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Workspace, echo=FALSE, message=FALSE, results='hide'}
#If need to clear 'Environment'
#First view environment if you like
#ls()
#To clear your environment
#remove(list = ls())

getwd()

# general
library(tidyverse)
library(readxl)
library(plyr)
library(dplyr)
library(devtools)
library(reshape2)
library(here)
"%notin%" = Negate('%in%')
here::here()
library(pander)

# graphics
library(ggplot2)
library(lattice)
library(graphics)
library(RColorBrewer)
library(gridExtra)
library(captioner)
library(viridis)
library("grid")
library("ggplotify")


#modelling
library(Hmisc)
library(mgcv)
library(unmarked)
library(lme4)
library(mvabund)
library(tweedie)
library(vegan)

## mvabund bug source code
source("code/2a_synthesis/traitglm.R")
source("code/2a_synthesis/get_polys.R")

```


## Models

**Manyglms for model comparisons**
  
- prob_life_spp (redone)

- prob_ocean_spp (redone)

- prob_lat_spp (redone)

- prob_life_clusters_many (redone)

- prob_ocean_clusters_many (redone)

- prob_life_trait_many (redone)

- prob_ocean_trait_many (redone)

- prob_life_phylo_many 

- prob_ocean_phylo_many

- prob_lat_phylo_many


**LASSO glm1path models for significance testing**
  
- prob_life_cluster_LASSO (redone)

- prob_ocean_cluster_LASSO (redone)

- prob_lat_cluster_LASSO

- prob_life_trait_LASSO (redone)

- prob_ocean_trait_LASSO (redone)

- prob_lat_trait_LASSO (redone)

- prob_life_phylo_LASSO

- prob_ocean_phylo_LASSO

- prob_lat_phylo_LASSO


### READ MODELS BACK IN

Species and environment matrix models - manyglm.

```{r READ IN SPP MODELS}
#/Users/tashhardy/Documents/GitHub/albacore-diet-global/outputs_figures/traitglms/prob_models

prob_life_spp = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_life_spp.rds")) 
prob_ocean_spp = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_ocean_spp.rds"))
prob_lat_spp = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_lat_spp.rds"))

```

Species, cluster number and environment models - manyglms + LASSOs.

```{r READ IN CLUSTER MODELS}

#Manyglms
prob_life_clusters_many = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_life_clusters_many.rds")) 
prob_ocean_clusters_many = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_ocean_clusters_many.rds"))
prob_lat_clusters_many = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_lat_clusters_many.rds"))

#LASSO glms
prob_life_cluster_LASSO = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_life_cluster_LASSO.rds"))
prob_ocean_cluster_LASSO = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_ocean_cluster_LASSO.rds"))
prob_lat_cluster_LASSO = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_lat_cluster_LASSO.rds"))

```

Species, phylogenetic info and environment models - manyglms + LASSOs

```{r READ IN PHYLO MODELS}

#Manyglms
prob_life_phylo_many = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_life_phylo_many.rds")) 
prob_ocean_phylo_many = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_ocean_phylo_many.rds"))
prob_lat_phylo_many = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_lat_phylo_many.rds"))

#LASSO glms
prob_life_phylo_LASSO = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_life_phylo_LASSO.rds"))
prob_ocean_phylo_LASSO = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_ocean_phylo_LASSO.rds"))
prob_lat_phylo_LASSO = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_lat_phylo_LASSO.rds"))

```


Species, traits and environment models - manyglms + LASSOs.

```{r READ IN TRAIT MODELS}

# Manyglms
prob_life_trait_many = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_life_trait_many.rds")) 
prob_ocean_trait_many = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_ocean_trait_many.rds"))
prob_lat_trait_many = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_lat_trait_many.rds"))

#LASSO glms
prob_life_trait_LASSO = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_life_trait_LASSO.rds"))
prob_ocean_trait_LASSO = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_ocean_trait_LASSO.rds"))
prob_lat_trait_LASSO = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_lat_trait_LASSO.rds"))

```

Interaction models

```{r READ IN INTERACTION MODELS}
#Manyglm
prob_inter_trait_many = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_inter_trait_many.rds"))
prob_inter_trait_many2 = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_inter_trait_many2.rds"))

#LASSO glms
prob_inter_trait_LASSO = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_inter_trait_LASSO.rds"))
prob_inter_trait_LASSO2 = readRDS(here("outputs_figures/traitglms/prob_models_sep2021/prob_inter_trait_LASSO2.rds"))

```

### MODEL COMPARISON

#### Model Outputs

Cluster, phylo, trait, spp models each for (i) predator life stage and (ii) ocean basins

##### Model: prob_life_spp

```{r Model Output prob_life_spp}
prob_life_spp
```

```{r Anova prob_life_spp}

#an_prob_life_spp = anova.traitglm(prob_life_spp)
#saveRDS(an_prob_life_spp, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_life_spp.rds"))
an_prob_life_spp = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_life_spp.rds"))
an_prob_life_spp

```

##### Model: prob_ocean_spp

```{r Model Output prob_ocean_spp}
prob_ocean_spp
```


```{r Anova prob_ocean_spp}

#an_prob_ocean_spp = anova.traitglm(prob_ocean_spp)
#saveRDS(an_prob_ocean_spp, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_ocean_spp.rds"))
an_prob_ocean_spp = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_ocean_spp.rds"))
an_prob_ocean_spp

```

##### Model: prob_lat_spp

```{r Model Output prob_lat_spp}
prob_lat_spp
```

```{r Anova prob_lat_spp}

#an_prob_lat_spp = anova.traitglm(prob_lat_spp)
#saveRDS(an_prob_lat_spp, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_lat_spp.rds"))
an_prob_lat_spp = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_lat_spp.rds"))
an_prob_lat_spp

```

##### Model: prob_life_clusters_many

```{r Model Output prob_life_clusters_many}
prob_life_clusters_many
```

```{r Anova prob_life_clusters_many}

#an_prob_life_clusters_many = anova.traitglm(prob_life_clusters_many)
#saveRDS(an_prob_life_clusters_many, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_life_clusters_many.rds"))
an_prob_life_clusters_many = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_life_clusters_many.rds"))
an_prob_life_clusters_many



```

##### Model: prob_ocean_clusters_many

```{r Model Output prob_ocean_clusters_many}
prob_ocean_clusters_many
```

```{r Anova prob_ocean_clusters_many}

#an_prob_ocean_clusters_many = anova.traitglm(prob_ocean_clusters_many)
#saveRDS(an_prob_ocean_clusters_many, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_ocean_clusters_many.rds"))
an_prob_ocean_clusters_many = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_ocean_clusters_many.rds"))
an_prob_ocean_clusters_many

```

##### Model: prob_lat_clusters_many

```{r Model Output prob_lat_clusters_many}
prob_lat_clusters_many
```

```{r Anova prob_lat_clusters_many}

#an_prob_lat_clusters_many = anova.traitglm(prob_lat_clusters_many)
#saveRDS(an_prob_lat_clusters_many, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_lat_clusters_many.rds"))
an_prob_lat_clusters_many = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_lat_clusters_many.rds"))
an_prob_lat_clusters_many

```

##### Model: prob_life_phylo_many

```{r Model Output prob_life_phylo_many}
prob_life_phylo_many
```

```{r Anova prob_life_phylo_many}

an_prob_life_phylo_many = anova.traitglm(prob_life_phylo_many)
saveRDS(an_prob_life_phylo_many, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_life_phylo_many.rds"))
an_prob_life_phylo_many = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_life_phylo_many.rds"))
an_prob_life_phylo_many

```

##### Model: prob_ocean_phylo_many

```{r Model Output prob_ocean_phylo_many}
prob_ocean_phylo_many
```

```{r Anova prob_ocean_phylo_many}

an_prob_ocean_phylo_many = anova.traitglm(prob_ocean_phylo_many)
saveRDS(an_prob_ocean_phylo_many, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_ocean_phylo_many.rds"))
an_prob_ocean_phylo_many = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_ocean_phylo_many.rds"))
an_prob_ocean_phylo_many

```

##### Model: prob_lat_phylo_many

```{r Model Output prob_lat_phylo_many}
prob_lat_phylo_many
```

```{r Anova prob_lat_phylo_many}

an_prob_lat_phylo_many = anova.traitglm(prob_lat_phylo_many)
saveRDS(an_prob_lat_phylo_many, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_lat_phylo_many.rds"))
an_prob_lat_phylo_many = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_lat_phylo_many.rds"))
an_prob_lat_phylo_many

```


##### Model: prob_life_trait_many

```{r Model Output prob_life_trait_many}
prob_life_trait_many
```

```{r Anova prob_life_trait_many}

#an_prob_life_trait_many = anova.traitglm(prob_life_trait_many)
#saveRDS(an_prob_life_trait_many, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_life_trait_many.rds"))
an_prob_life_trait_many = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_life_trait_many.rds"))
an_prob_life_trait_many

```

##### Model: prob_ocean_trait_many

```{r Model Output prob_ocean_trait_many}
prob_ocean_trait_many
```


```{r Anova prob_ocean_trait_many}

#an_prob_ocean_trait_many = anova.traitglm(prob_ocean_trait_many)
#saveRDS(an_prob_ocean_trait_many, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_ocean_trait_many.rds"))
an_prob_ocean_trait_many = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_ocean_trait_many.rds"))
an_prob_ocean_trait_many

```

##### Model: prob_lat_trait_many

```{r Model Output prob_lat_trait_many}
prob_lat_trait_many
```

```{r Anova prob_lat_trait_many}

#an_prob_lat_trait_many = anova.traitglm(prob_lat_trait_many)
#saveRDS(an_prob_lat_trait_many, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_lat_trait_many.rds"))
an_prob_lat_trait_many = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_lat_trait_many.rds"))
an_prob_lat_trait_many

```

##### Model: prob_inter_trait_many

```{r Model Output: prob_inter_trait_many}
prob_inter_trait_many

prob_inter_trait_many$aic

#prob_inter_trait_many$two.loglike same as deviance explained

prob_inter_trait_many$deviance

prob_inter_trait_many$call

#temp = as.data.frame(prob_inter_trait_many$coefficients)


```

```{r Anova prob_inter_trait_many}

#an_prob_inter_trait_many = anova.traitglm(prob_inter_trait_many)
#saveRDS(an_prob_inter_trait_many, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_inter_trait_many.rds"))
an_prob_inter_trait_many = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_inter_trait_many.rds"))
an_prob_inter_trait_many

```

##### Model: prob_inter_trait_many2

```{r Model Output: prob_inter_trait_many2}

prob_inter_trait_many2

```

```{r Anova prob_inter_trait_many2}

#an_prob_inter_trait_many2 = anova.traitglm(prob_inter_trait_many2)
#saveRDS(an_prob_inter_trait_many2, file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_inter_trait_many2.rds"))
an_prob_inter_trait_many2 = readRDS(file = here("outputs_figures/traitglms/prob_models_sep2021/an_prob_inter_trait_many2.rds"))
an_prob_inter_trait_many2

```

## Conclusions

The species models had the lowest residual deviance explained, and typically higher AIC values.
Cluster models performed slightly better than traits models, but not massive difference. Both cluster and trait sets of models outperformed species models.

The best model in terms of residual deviance explained was the latitude + traits model, however, did have higher AIC scores.
The interaction model that included ocean basin + life stage had the lowest AIC but low deviance explained.

## Remaining to do's

- Run a null model?
  - Run latitude + cluster model?
  
## Model Comp Outputs

### Model List
```{r Model list for outputs, eval=FALSE}

#Species*env models
prob_life_spp
prob_ocean_spp
prob_lat_spp

#Cluster*env models (manyglm and LASSO)
#Manyglms
prob_life_clusters_many
prob_ocean_clusters_many
prob_lat_clusters_many
#LASSO glms
prob_life_cluster_LASSO
prob_ocean_cluster_LASSO
prob_lat_cluster_LASSO

#Phylo*env models (manyglm and LASSO)
#Manyglms
prob_life_phylo_many
prob_ocean_phylo_many
prob_lat_phylo_many
#LASSO glms
prob_life_phylo_LASSO
prob_ocean_phylo_LASSO
prob_lat_phylo_LASSO

#Trait*env models
# Manyglms
prob_life_trait_many 
prob_ocean_trait_many
prob_lat_trait_many
#LASSO glms
prob_life_trait_LASSO
prob_ocean_trait_LASSO
prob_lat_trait_LASSO 

#Additive multi-variable env models
#Manyglm
prob_inter_trait_many
prob_inter_trait_many2
#LASSO glms
prob_inter_trait_LASSO
prob_inter_trait_LASSO2
```

**Note**
  
  - Could add additive models, LASSO models?
  
### Model outputs without phylo
  
```{r Make output dataframe}

model = c("prob_life_spp", "prob_ocean_spp", "prob_lat_spp", 
          "prob_life_trait_many", "prob_ocean_trait_many", "prob_lat_trait_many",
          "prob_life_clusters_many", "prob_ocean_clusters_many", "prob_lat_clusters_many") 
#,        "prob_inter_trait_many", "prob_inter_trait_many2"
mod_type = c("species", "species", "species", 
             "traits", "traits", "traits",
             "cluster", "cluster", "cluster")
#,           "trait*ocean*life stage", "trait*ocean*climate"
covar = c("predator life stage", "ocean basin", "climate",
          "predator life stage", "ocean basin", "climate",
          "predator life stage", "ocean basin", "climate")
#,        "ocean*life stage", "ocean*climate"
deviance = c(prob_life_spp$deviance, prob_ocean_spp$deviance, prob_lat_spp$deviance,
             prob_life_trait_many$deviance, prob_ocean_trait_many$deviance, prob_lat_trait_many$deviance,
             prob_life_clusters_many$deviance, prob_ocean_clusters_many$deviance, prob_lat_clusters_many$deviance)
#,           prob_inter_trait_many$deviance, prob_inter_trait_many2$deviance
aic = c(prob_life_spp$AICsum, prob_ocean_spp$AICsum, prob_lat_spp$AICsum,
        prob_life_trait_many$AICsum, prob_ocean_trait_many$AICsum, prob_lat_trait_many$AICsum,
        prob_life_clusters_many$AICsum, prob_ocean_clusters_many$AICsum, prob_lat_clusters_many$AICsum)
#,      prob_inter_trait_many$AICsum, prob_inter_trait_many2$AICsum
mod_dev = data.frame(covar, mod_type, model, deviance, aic)

write.csv(mod_dev, here("data/output_data/mod_dev_sep2021.csv"))
```

```{r Make output graph}
mod_dev$mod_type = as.factor(mod_dev$mod_type)
mod_dev$mod_type = factor(mod_dev$mod_type, levels = c("species", "traits", "cluster"))
levels(mod_dev$mod_type)

mod_dev_graph = ggplot(data = mod_dev, aes(x = mod_type, y = deviance, fill = mod_type)) + 
  geom_bar(stat = "identity", width = 1) + 
  theme_classic() +
  theme(axis.title = element_text(size=18),
        axis.text = element_text(size=18),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size=18),
        legend.title = element_text(size=18),
        legend.position = "top") +
  xlab("Covariate fitted to model") + ylab("Deviance explained") + 
  scale_fill_viridis_d(option="D", begin = 0.2, end = 0.9, name = "Model type")+
  facet_grid(.~ covar) +
  theme(strip.text.x = element_text(size = 18))

mod_dev_graph

mod_dev_graph2 = ggplot(data = mod_dev) + 
  geom_bar(aes(y = deviance, fill = mod_type)) + #, #x= mod_type, 
           #position = position_dodge(0.5), size=2) +
  #geom_bar(width=0.9, position=position_dodge())+
  theme_classic() +
  theme(axis.title = element_text(size=18),
        axis.text = element_text(size=18),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size=18),
        legend.title = element_text(size=18),
        legend.position = "top") +
  xlab("Covariate fitted to model") + ylab("Deviance explained") + 
  scale_fill_viridis_d(option="D", begin = 0.2, end = 0.9, name = "Model type")+
  facet_grid(.~ covar) +
  theme(strip.text.x = element_text(size = 18))

mod_dev_graph2

ggsave(here("outputs_figures/traitglms/prob_models_sep2021/mod_dev_graph.pdf"), 
       plot = mod_dev_graph, width=8, height=4, dpi=300)

ggsave(here("outputs_figures/traitglms/prob_models_sep2021/mod_dev_graph.jpeg"), 
       plot = mod_dev_graph, width=8, height=4, dpi=300)

```

### Model outputs with phylo

```{r Make phylo output dataframe}

model = c("prob_life_spp", "prob_ocean_spp", "prob_lat_spp", 
          "prob_life_trait_many", "prob_ocean_trait_many", "prob_lat_trait_many",
          "prob_life_clusters_many", "prob_ocean_clusters_many", "prob_lat_clusters_many",
          "prob_life_phylo_many", "prob_ocean_phylo_many", "prob_lat_phylo_many") 
#,        "prob_inter_trait_many", "prob_inter_trait_many2"
mod_type = c("species", "species", "species", 
             "traits", "traits", "traits",
             "cluster", "cluster", "cluster",
             "phylo", "phylo", "phylo")
#,           "trait*ocean*life stage", "trait*ocean*climate"
covar = c("predator life stage", "ocean basin", "climate",
          "predator life stage", "ocean basin", "climate",
          "predator life stage", "ocean basin", "climate",
          "predator life stage", "ocean basin", "climate")
#,        "ocean*life stage", "ocean*climate"
deviance = c(prob_life_spp$deviance, prob_ocean_spp$deviance, prob_lat_spp$deviance,
             prob_life_trait_many$deviance, prob_ocean_trait_many$deviance, prob_lat_trait_many$deviance,
             prob_life_clusters_many$deviance, prob_ocean_clusters_many$deviance, prob_lat_clusters_many$deviance,
             prob_life_phylo_many$deviance, prob_ocean_phylo_many$deviance, prob_lat_phylo_many$deviance)
#,           prob_inter_trait_many$deviance, prob_inter_trait_many2$deviance
aic = c(prob_life_spp$AICsum, prob_ocean_spp$AICsum, prob_lat_spp$AICsum,
        prob_life_trait_many$AICsum, prob_ocean_trait_many$AICsum, prob_lat_trait_many$AICsum,
        prob_life_clusters_many$AICsum, prob_ocean_clusters_many$AICsum, prob_lat_clusters_many$AICsum,
        prob_life_phylo_many$AICsum, prob_ocean_phylo_many$AICsum, prob_lat_phylo_many$AICsum)
#,      prob_inter_trait_many$AICsum, prob_inter_trait_many2$AICsum
mod_dev_phylo = data.frame(covar, mod_type, model, deviance, aic)

write.csv(mod_dev_phylo, here("data/output_data/mod_dev_phylo_sep2021.csv"))
```

```{r Make phylo output graph}
#mod_dev_phylo$mod_type = as.factor(mod_dev_phylo$mod_type)
mod_dev_phylo$mod_type = factor(mod_dev_phylo$mod_type, levels = c("species", "phylo", "traits", "cluster"))
levels(mod_dev_phylo$mod_type)

#mod_dev_graph = ggplot(data = mod_dev, aes(x = model, y = deviance, color = covar)) + 
#  geom_bar(stat = "identity", width = 1) + 
#  facet_grid(.~ mod_type)

mod_dev_graph_phy = ggplot(data = mod_dev_phylo, aes(x = mod_type, y = deviance, fill = mod_type)) + 
  geom_bar(stat = "identity", width = 1) + #ylim(2000, 4100) +
  theme_classic() +
  theme(axis.title = element_text(size=18),
        axis.text = element_text(size=18),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size=18),
        legend.title = element_text(size=18),
        legend.position = "top") +
  xlab("Covariate fitted to model") + ylab("Deviance explained") + 
  #scale_y_continuous(limits=c(2000, 4200), breaks = NULL) +
  scale_fill_viridis_d(option="D", begin = 0.2, end = 0.9, name = "Model type")+
  facet_grid(.~ covar) +
  theme(strip.text.x = element_text(size = 18))
mod_dev_graph_phy

mod_dev_graph_phy2 = ggplot(data = mod_dev_phylo) + 
  geom_bar(aes(y = deviance, fill = mod_type)) +
  theme_classic() +
  theme(axis.title = element_text(size=18),
        axis.text = element_text(size=18),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size=18),
        legend.title = element_text(size=18),
        legend.position = "top") +
  xlab("Covariate fitted to model") + ylab("Deviance explained") + 
  scale_fill_viridis_d(option="D", begin = 0.2, end = 0.9, name = "Model type")+
  facet_grid(.~ covar) +
  theme(strip.text.x = element_text(size = 18))

mod_dev_graph_phy2

ggsave(here("outputs_figures/traitglms/prob_models_sep2021/mod_dev_phylo_graph.pdf"), 
       plot = mod_dev_graph_phy, width=8, height=4, dpi=300)

ggsave(here("outputs_figures/traitglms/prob_models_sep2021/mod_dev_phylo_graph.jpeg"), 
       plot = mod_dev_graph_phy, width=8, height=4, dpi=300)

ggsave(here("outputs_figures/traitglms/prob_models_sep2021/mod_dev_phylo_graph2.pdf"), 
       plot = mod_dev_graph_phy2, width=8, height=4, dpi=300)

ggsave(here("outputs_figures/traitglms/prob_models_sep2021/mod_dev_phylo_graph2.jpeg"), 
       plot = mod_dev_graph_phy2, width=8, height=4, dpi=300)

```
