---
title: "Step 1 Data Manip"
author: "Natasha Hardy"
date: "October 10, 2019"
output:
  html_document: default
  pdf_document: default
---

## About

This document contains code for manipulating trait information for 5 prey traits, and life history estimation, for albacore prey species from our global meta-analysis of albacore diet composition.

From the MS:

"2.2 Prey trait information

For each prey species and main life stage distinctions (i.e., larva, juvenile, adult), we collected trait information for four habitat use traits and one behavioural trait known to affect the likelihood of a pelagic predator encountering, attacking and capturing prey (Green et al., 2019). These were: (i) vertical and (ii) horizontal habitat association, (iii) presence of diel vertical migration, (iv) presence of seasonal migration and (v) aggregation behaviour (Supplementary Data, Table S4 & S5). For analyses, we include the phylogenetic Order verified using Open Tree of Life Data (OpenTree et al, 2020). Trait data were compiled for a broader database construction effort of traits that inform predator-prey interactions for albacore (Gleiber et al. in prep). We further describe how prey species’ trait values were collected in Appendix B (Supporting Information)."

## Workpsace

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r Load Workspace, message=FALSE}

#R documentation
library(pander)
# general
library(tidyverse)
library(readxl)
library(readr)
library(plyr)
library(dplyr)
library(devtools)
library(here)
"%notin%" = Negate('%in%')
here::here()

```

## Outputs

List of outputs created for analyses:

- Prey species' coarse trait information *data/output_data/prey_trait_select.csv*

- Albacore diet data *data/output_data/alb_global_diet_total.csv*

- Albacore diet sampling covariates *alb_covars.csv*


## Dataframes

### Diet raw data

Load diet data and merge with prey size trait information, and undertake necessary data selection and manipulation of variables.

```{r Albacore diet data load & manipulation, tidy=TRUE}

#Load diet data spreadsheet
alb_global = as.data.frame(read_xlsx(here::here("data/input_data/albacore_diet_global.xlsx"),
                      sheet = "albacore_diet_global")) %>% #2134 rows and 77 vars
  dplyr::filter(TaxLev == "species") %>% #1107 rows and 77 vars
  #There are ~1105 observations when we filter by species
  dplyr::filter(Include %in% c("Yes", "Maybe") & !pred_life == "larvae") %>% #1028 rows
  #1028 when we remove larval albacore diet studies (there are 3)
  #dplyr::rename() %>%
  dplyr::select(StudyID, CiteAuth, CiteYear, OceanBasin, OceanBasinQ, LocatName,
                LocatLatitude:LocatLongitude, MonthRange, YearEnd, stomachs_used,
                pred_life:est_ref, stomachs_used, prey_class:prey_sp,
                prey_age_reported_1:prey_age_use, Include, FO, 
                pFOuse, N, pN, `Total Mass (g)`, pM, maxl_use, maxl_type) %>%
  mutate(gape_lmax = pred_flmean*0.0823+1.758, 
         gape_height_limit = pred_flmean*0.0246+4.603,
         prey_age_reported_1=case_when(
                             #prey_age_reported_1 == "NA" ~ "none",
                             prey_age_reported_1 == "larvae" ~ "larva",
                             prey_age_reported_1 == "juvenile" ~ "juvenile",
                             prey_age_reported_1 == "adult" ~ "adult"))
  #1028 obsm 44 vars
  #calculates the gape length and height limits
  #Reference paper for this calculation -- Menard et al. (2006) for yellowfin tuna

#Correct data type
alb_global$maxl_use <- as.numeric(alb_global$maxl_use)
#The NA's in the reported age for prey column cause problems later and need 
#to be assigned a categorical value
alb_global$prey_age_reported_1 <- as.character(alb_global$prey_age_reported_1)
alb_global$prey_age_reported_1 <- replace_na(alb_global$prey_age_reported_1, "none")
#check this
unique(as.factor(alb_global$prey_age_reported_1))

#Change vector type for YearEnd variable
alb_global$YearEnd <- as.integer(alb_global$YearEnd)

#Deal with NA's in stomachs used column
alb_global$stomachs_used <- as.integer(alb_global$stomachs_used)
alb_global$stomachs_used <- as.character(alb_global$stomachs_used)
alb_global$stomachs_used <- replace_na(alb_global$stomachs_used, "not reported")
unique(alb_global$stomachs_used)
unique(alb_global$prey_sp)

#Save total df version
write.csv(alb_global, here::here("data/output_data/alb_global_diet.csv"), 
          row.names = FALSE) 
#1028 observations and 41 variables

```


### Covariates

Note that we lose two replicates here:

- McHugh 1952 (Mexico to USA) due to containing only higher level taxa classifications and not species-level data

- Williams 2015 (NZ_L_2010) due to data entry error or copy/paste error, data have been copied over.


```{r Covariates, tidy = TRUE}

alb_covars = alb_global %>%
  dplyr::select(StudyID, CiteYear, OceanBasin, OceanBasinQ, LocatName, LocatLatitude:pred_life_est) %>% 
  #, stomachs_used
  distinct(StudyID, CiteYear, OceanBasin, OceanBasinQ, LocatName, LocatLatitude, 
           LocatLongitude, YearEnd, MonthRange, stomachs_used, pred_life) %>% 
  #, stomachs_used
  mutate(lat_cat = if_else(abs(LocatLatitude) > 23.43662, "temperate", "tropical"),
         pred_life_adjust=case_when(pred_life == "juvenile" ~ "juvenile",
                             pred_life == "juvenile, adult" ~ "mixed",
                             pred_life == "none" ~ "mixed",
                             pred_life == "adult" ~ "adult"))

alb_covars$rep_ID = paste(alb_covars$StudyID, row.names(alb_covars), sep = "_")

#Merge these back to the diet df
alb_global_diet = alb_global %>%
  left_join(alb_covars)

write.csv(alb_covars, here::here("data/output_data/alb_covars.csv"), row.names = FALSE)

```

## Data checks & additional info

### Prey length info reported

```{r Reported prey length information, tidy = TRUE}

alb_global_prey_lengths = alb_global %>%
  filter(maxl_use > 0) %>%
  dplyr::select(prey_class:prey_sp, pFOuse, pN, pM, maxl_use, maxl_type) %>%
  group_by(prey_sp, maxl_type) %>% #, life_stage , pFOuse, pN, pM
  dplyr::summarize(maxFO = max(pFOuse, na.rm=TRUE), 
                   maxN = max(pN, na.rm=TRUE), 
                   maxM = max(pM, na.rm=TRUE), 
                   maxL = max(maxl_use, na.rm=TRUE)) 

write.csv(alb_global_prey_lengths, here::here("data/output_supp/alb_global_prey_lengths.csv"), row.names = FALSE)

```

### Prey life stage reported

```{r Reported prey lifestage information, tidy = TRUE}

alb_global_prey_life = alb_global %>%
  filter(prey_age_reported_1 != "none",) %>%
  dplyr::select(prey_class:prey_sp, prey_age_reported_1, pFOuse, pN, pM, maxl_use, maxl_type) %>%
  group_by(prey_sp, prey_age_reported_1, maxl_type) %>% #, life_stage , pFOuse, pN, pM
  dplyr::summarize(maxFO = max(pFOuse, na.rm=TRUE), 
                   maxN = max(pN, na.rm=TRUE), 
                   maxM = max(pM, na.rm=TRUE), 
                   maxL = max(maxl_use, na.rm=TRUE)) 

write.csv(alb_global_prey_life, here::here("data/output_supp/alb_global_prey_life.csv"), row.names = FALSE)

```

### Prey prevalence data reported

```{r Reported diet metrics summary v1, tidy = TRUE}

alb_global_metrics = alb_global %>%
  #unique(c(alb_global_prob$prey_sp, alb_global_prob$prey_select))
  dplyr::select(prey_class:prey_sp, prey_age_reported_1, pFOuse, pN, pM, maxl_use, maxl_type) %>%
  group_by(prey_sp, prey_age_reported_1, maxl_type) %>% #, life_stage , pFOuse, pN, pM
  dplyr::summarize(maxFO = max(pFOuse, na.rm=TRUE), 
                   maxN = max(pN, na.rm=TRUE), 
                   maxM = max(pM, na.rm=TRUE), 
                   maxL = max(maxl_use, na.rm=TRUE)) 

write.csv(alb_global_metrics, here::here("data/output_supp/alb_global_metrics.csv"), row.names = FALSE)

```

```{r Reported diet metrics summary v2, tidy = TRUE}

alb_global_metrics_basic = alb_global %>%
  #unique(c(alb_global_prob$prey_sp, alb_global_prob$prey_select))
  dplyr::select(prey_class:prey_sp, pFOuse, pN, pM) %>%
  group_by(prey_sp) %>% #, life_stage , pFOuse, pN, pM
  dplyr::summarize(maxFO = max(pFOuse, na.rm=TRUE), 
                   maxN = max(pN, na.rm=TRUE), 
                   maxM = max(pM, na.rm=TRUE)) 

write.csv(alb_global_metrics_basic, here::here("data/output_supp/alb_global_metrics_basic.csv"), row.names = FALSE)

```



## Prey trait db

### Load and select traits for analyses

Loading in and selecting key species coarse traits: for habitat association, seasonal, vertical migration and aggregation behaviour. Note that NH overhauled our prey life stage estimation protocol in December 2021 through early 2022, and we have thoroughly curated this step.

```{r Prey trait data load, message=FALSE, results='hide', tidy=TRUE}

#Load prey trait db
prey_trait <- as.data.frame(read_xlsx(here::here("data/input_data/prey_trait_db.xlsx"),
                                           sheet = "prey_trait_db"))

```

Key traits selection:

```{r Prey traits selection, tidy=TRUE}

prey_trait_select = prey_trait %>%
  dplyr::select(prey_class:prey_sp, life_stage, diet_likely, life_note, 
                vert_habitat, horz_habitat, diel_migrant, diel_migrant_cat, refuge_cat, season_migrant, season_cat, gregarious_primary) %>%
  filter(life_stage == diet_likely)

prey_trait_select$prey_sp[duplicated(prey_trait_select$prey_sp)]

str(prey_trait_select)

```

Note that there are 308 spp remaining in the global diet db. Why the difference in 5 species from the trait db? These were the prey taxa found in the larval albacore tuna diets and were thus previously removed.

```{r check why some species drop out}

check_spp = prey_trait_select %>%
  anti_join(alb_global_metrics_basic, by="prey_sp")

```

```{r filter prey trait df by species remaining in diet df}

prey_trait_select_use = prey_trait_select %>%
  filter(prey_sp %in% unique(alb_global_metrics_basic$prey_sp)) %>%
  left_join(alb_global_metrics_basic)

write.csv(prey_trait_select_use, here::here("data/output_data/prey_trait_select.csv"), row.names = FALSE)

```

## GO TO 2_extract_longhurst to work on Figure 1a
## THEN TO document 3a_alb_data_covars to work on covariate info
## THEN TO 4_species_accum doc to work on Figure 1b and c
## THEN TO 5_trait_phylo_graph for species and trait phylogenetic graphs
