---
title: "Step 4 Species Accumulation"
author: "Natasha Hardy"
date: "21/03/2022"
output: html_document
---
---
title: "Albacore Diet Synthesis Rarefaction"
author: "Natasha Hardy"
date: "20/10/2021"
output: html_document
---

## About

Illustrating the biological diversity in diets of a highly migratory predator

## Workspace

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r Load Workspace, message=FALSE}

#R documentation
library(pander)
# general
library(tidyverse)
library(readxl)
library(readr)
library(plyr)
library(dplyr)
library(devtools)
library(here)
"%notin%" = Negate('%in%')
here::here()
# RE-INSTALLED THESE

#Graphics
library(ggplot2)
library(viridis)
library(RColorBrewer)
library(wesanderson)

#Biodiversity
library(BiodiversityR)

library(vegan)
# From GitHub
#install_github("MoBiodiv/mobr")
library(mobr)
#devtools::install_github('bbc/bbplot')
library(bbplot)
```

## Rarefaction data

```{r Rarefaction curves data}
#Data -- using total diet data
alb_global_dietpa = read.csv(here("data/output_data/alb_global_wide_dietpa.csv"), 
          check.names = FALSE) %>%
  group_by(OceanBasin) %>%
  arrange(YearEnd)

## Albacore prey species P/A data
# data
PreyPAdata = as.data.frame(alb_global_dietpa[,10:ncol(alb_global_dietpa)])
PreyPAsite = as.data.frame(alb_global_dietpa[,c(1:5)])

PreyPAsite$YearEnd = as.integer(PreyPAsite$YearEnd)
PreyPAsite$StudyID = as.factor(PreyPAsite$StudyID)
PreyPAsite$OceanBasin = as.factor(PreyPAsite$OceanBasin)

str(PreyPAdata)
str(PreyPAsite)
View(PreyPAsite)

rownames(PreyPAsite) = rownames(PreyPAdata)
#check.datasets(PreyPAdata, PreyPAsite) #ensure that the sites of the community dataset are the same sites as those from the environmental dataset --> something that is assumed to be the case for the BiodiversityR and vegan packages.


write.csv(PreyPAsite, here("data/output_data/alb_global_rarefactiondata.csv"), row.names = FALSE)

```

## Basic rarefaction curves for species-level taxa:

```{r Summary species specnumber}

# number of species and the factor that I want to group them by
SpO <- specnumber(PreyPAdata, PreyPAsite$OceanBasin)
SpO

```

```{r Summary rarefaction and accum result}

# Acummresult function

#Produces average species accumulation curve
Alb.accum <- accumresult(PreyPAdata, y=PreyPAsite$OceanBasin, method='exact', conditioned=TRUE)
#str(Alb.accum)
accumplot(Alb.accum)

#Produces cumulative species curves
Alb.accum2 <- accumresult(PreyPAdata, y=PreyPAsite$OceanBasin, method='collector', conditioned=TRUE)
#str(Alb.accum2)
accumplot(Alb.accum2)

```

## Variable rarefaction by Ocean Basin

```{r Rarefaction by sites - exact method, eval=FALSE}

#Produces average species accumulation curve
#Plot <- accumcomp(PreyPAdata, y=PreyPAsite, factor='OceanBasin',
#                  method='exact', conditioned=TRUE, plotit=TRUE) #scale = 'YearEnd'

#Plot.long <- accumcomp.long(Plot)
#str(Plot.long)

```


```{r Rarefaction by sites - collector method}
#Plot.long.covar = Plot.long %>%
#  rename(OceanBasin = Grouping) %>%
#  left_join(PreyPAsite, by = "OceanBasin")

#Produces cumulative species curves
Plot2 <- accumcomp(PreyPAdata, y=PreyPAsite, factor='OceanBasin', 
                   #scale = 'YearEnd', 
                   method='collector', conditioned=TRUE, plotit=TRUE)

Plot2.long <- accumcomp.long(Plot2)
#str(Plot2.long)
# save and export the accumcomp data in order to add SW back in with just the one point

# Did this for first and second Plot output for accumcomp function.
#write.csv(Plot.long, here("data/output_data/accum_plot.csv"), row.names = FALSE)
#accum.data <- read.csv(here("data/output_data/accum_plot.csv"))
#View(accum.data)

#write.csv(Plot2.long, here("data/output_data/accum2_plot.csv"), row.names = FALSE)
#accum2.data <- read.csv(here("data/output_data/accum2_plot.csv"))
#View(accum2.data)
```

Problem: Still can't get the data for year to append to the rarefaction plot data.
Solution: find a way to append the YearEnd values, seeing as those data were ordered and the fed to the rarefaction code routine...

```{r Manipulate and extract year data}

prey_year = PreyPAsite %>%
  arrange(desc(OceanBasin))

plot2_data = cbind(prey_year, Plot2.long)

```

```{r Loop to extract year data for plot 1}

Plot.long$Year <- NA
groupings <- unique(Plot.long$Grouping)
 
for(g in 1:length(groupings)) {
  group <- groupings[g]
  # assuming Grouping was based on variable Region of the env.data
  Plot.long[Plot.long$Grouping == group, "Year"] <- PreyPAsite[PreyPAsite$OceanBasin == group, "YearEnd"]
  }

```


```{r rarefaction graphs - curve, eval=FALSE}

levels(as.factor(Plot.long$Grouping))

Plot.long$Grouping <- factor(Plot.long$Grouping, levels = c("Mediterranean", "N Atlantic",
                                                        "S Atlantic", "Indian",
                                                        "N Pacific", "S Pacific"))

Accum.plot <- ggplot(data = Plot.long, aes(y=Richness, x=Sites, colour=Grouping)) +
  geom_point(size=2) +
  geom_line(size=1.5) + 
  theme_classic() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  #scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15))+
  theme(axis.title=element_text(size=18)) +
  theme(axis.text=element_text(size=18)) +
  theme(legend.text=element_text(size=18)) +
  theme(legend.title=element_text(size=18)) +
  labs(x="Years sampled", y="Species cumulative richness")+
  scale_colour_manual(values=wes_palette("Zissou1", 7, type="continuous"), name="Ocean Basin")+
  geom_errorbar(data = Plot.long, aes(ymin=Richness-SD, ymax=Richness+SD), width=.2)
Accum.plot

ggsave(here("outputs_figures/accum.plot1.jpeg"), plot=Accum.plot, height = 8, width = 12, dpi = 300)
```


```{r Loop from Roeland for plot 2}

Plot2.long$Year <- NA
groupings <- unique(Plot2.long$Grouping)
 
for(g in 1:length(groupings)) {
  group <- groupings[g]
  # assuming Grouping was based on variable Region of the env.data
  Plot2.long[Plot2.long$Grouping == group, "Year"] <- PreyPAsite[PreyPAsite$OceanBasin == group, "YearEnd"]
  }

```

```{r rarefaction graphs - cumulative}
##Second plot
str(Plot2.long)

levels(as.factor(Plot2.long$Grouping))

Plot2.long$Grouping <- factor(Plot2.long$Grouping, levels = c("Mediterranean", "N Atlantic",
                                                        "S Atlantic", "Indian",
                                                        "N Pacific", "S Pacific"))

Accum.plot2 <- ggplot(data = Plot2.long, aes(y=Richness, x=Year, colour=Grouping)) +
  geom_point(size=2) +
  geom_line(size=1.5) + 
  theme_classic() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  #scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15))+
  #geom_text(aes(x=Year, label=Stomachs), size=4, hjust=0.5, col="black") +
  #geom_text_repel(data = accum.data2, aes(x=Year, label=Stomachs),
  #                nudge_x = -10, nudge_y = 10) +
  theme(axis.title=element_text(size=18)) +
  theme(axis.text=element_text(size=18)) +
  theme(legend.text=element_text(size=18)) +
  theme(legend.title=element_text(size=18)) +
  labs(x="Years sampled", y="Species cumulative richness")+
  #scale_colour_manual(values=wes_palette("Zissou1", 7, type="continuous"), name="Ocean Basin")
  scale_colour_manual(values=pnq_palette("Zissou1", 7, type="continuous"), name="Ocean Basin")
Accum.plot2

ggsave(here("outputs_figures/accum.plot2.jpeg"), plot=Accum.plot2, height = 8, width = 12, dpi = 300)


#Note no error bar for pure accumulation plot

```

## Comms with Roeland Kindt

Dear Natasha,
 
Thank you very much for email and finding BiodiversityR useful.
 
To answer briefly (on holiday in Mombasa right now)
 
The accumcomp function does not include information on characteristics of sites. This is because typical methods randomize the sites. But indeed for the collector method, the first site is the first site that is encountered
Therefore indeed this information needs to be added, maybe using a script similar to the following after getting data first in the accumcomp.long format
 
accum.long1$Year <- NA
groupings <- unique(accum.long1$Grouping)
 
for (g in 1:length(groupings) {
group <- groupings[g]
Assuming Grouping was based on variable Region of the env.data
accum.long1[accum.long1$Grouping == group, “Year”] <- env.data[env.data$Region == group, “Year”]
}
 
Then it should be possible to use ggplot as in your example, with year as x variable. It should be a numeric variable to generate the type of plot that you made earlier.
 
Please come back to me in case I misunderstood something.
 
Best regards – stay safe,
 
Roeland
 
Roeland KINDT, dr. ir.