---
title: "Albacore Review Diet Data Manipulation"
author: "Natasha Hardy"
date: "03/08/2021"
output: html_document
---

## About

This document contains code for manipulating diet data sets for albacore prey species from our global meta-analysis of albacore diet composition.

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r Load Workspace, message=FALSE}

#R documentation
library(pander)
# general
library(tidyverse)
library(readxl)
library(readr)
library(plyr)
library(dplyr)
library(devtools)
library(here)
"%notin%" = Negate('%in%')
here::here()

```


## Outputs needed for these manips:

- Prey species' morphometric trait information (body shape, eye diameter to body, and standard to total length): *data/output_data/prey_morph_sum.csv*

- Prey species' nutritional composition traits (energy density, percent lipid and percent protein composition): *data/output_data/prey_qual_sum.csv*

- Prey species' trait values: (i) selected priority trait columns *data/output_data/prey_trait_select.csv* and (ii) trait values for adult life stages only *data/output_data/prey_trait_adult.csv* --> for downstream data joins and analyses.

- Prey species length ratios for max and min length: *data/output_data/prey_length_wide.csv* 

- Prey species + filled in length ratios: *data/output_data/prey_length_ratiosp.csv*

## Albacore diet review data

Here we need to load up the albacore global diet data to clean and select what we want.
Tasks performed in these chunks:

- Select prey probable life stage;

- Merge probable prey life stage with appropriate traits;

- Retain frequency of occurrence + stomach sampled information for traitglm.

```{r Albacore diet data load & manipulation}

#Load diet data spreadsheet
#need this: prey_length_ratiosp
#This contains species as rows for which we have useable species through to class-level ratio information on larval:adult and juvenile:adult lengths

#Also need prey_length_wide data for l_max for adult column
#Should be able to get this from prey_length_ratiosp

alb_global = as.data.frame(read_xlsx(here("data/input_data/albacore_diet_global.xlsx"),
                      sheet = "albacore_diet_global")) %>% #2134 rows and 77 vars
  dplyr::filter(TaxLev == "species") %>% #1107 rows and 77 vars
  #There are ~1105 observations when we filter by species
  dplyr::filter(Include %in% c("Yes", "Maybe") & !pred_life == "larvae") %>% #1028 rows
  #1028 when we remove larval albacore diet studies (there are 3)
  #dplyr::rename() %>%
  dplyr::select(StudyID, CiteAuth, CiteYear, OceanBasin, OceanBasinQ, LocatName, LocatLatitude:LocatLongitude, MonthRange, YearEnd, stomachs_used, pred_life:est_ref, stomachs_used, prey_class:prey_sp, prey_age_reported_1:prey_age_use, Include, FO, pFOuse, N, pN, `Total Mass (g)`, pM, maxl_use, maxl_type) %>%
  mutate(gape_lmax = pred_flmean*0.0823+1.758, gape_height_limit = pred_flmean*0.0246+4.603) %>% #1028 obsm 44 vars
  #calculates the gape length and height limits
  #Add reference paper for this calculation -- Menard et al. (2006) for yellowfin tuna
  left_join(prey_morph_sum, by="prey_sp") #added morph info --> 1028 obs, 45 vars

#Correct data type
alb_global$maxl_use <- as.numeric(alb_global$maxl_use)

#Cleaning metrics
alb_global = alb_global %>% 
  mutate(maxtl_use = if_else(alb_global$maxl_type %in% c("SL", "ML"), maxl_use/standard_total, maxl_use),
         prey_age_reported_1=case_when(
                             #prey_age_reported_1 == "NA" ~ "none",
                             prey_age_reported_1 == "larvae" ~ "larva",
                             prey_age_reported_1 == "juvenile" ~ "juvenile",
                             prey_age_reported_1 == "adult" ~ "adult")
         ) #%>%

#The NA's in the reported age for prey column cause problems later and need to be assigned a categorical value
alb_global$prey_age_reported_1 <- as.character(alb_global$prey_age_reported_1)
alb_global$prey_age_reported_1 <- replace_na(alb_global$prey_age_reported_1, "none")

#Save total df
write.csv(alb_global, here("data/output_data/alb_global_diet_total.csv")) 
#1028 observations and 43 variables

```

The following summarises the max observed FO, N and M for each species in the dataset. This information can be useful when investigating the relative importance of species in albacore diets.

```{r Alb adult diet metrics summary}

alb_global_ad_metrics = alb_global %>%
  #unique(c(alb_global_prob$prey_sp, alb_global_prob$prey_select))
  dplyr::select(prey_class:prey_sp, pFOuse, pN, pM) %>%
  group_by(prey_sp) %>% #, life_stage , pFOuse, pN, pM
  dplyr::summarize(maxFO = max(pFOuse, na.rm=TRUE), maxN = max(pN, na.rm=TRUE), maxM = max(pM, na.rm=TRUE)) 

write.csv(alb_global_ad_metrics, here("data/output_data/alb_global_ad_metrics"))
```

Here we want to estimate the probable life stage of the species consumed based on their size ratios and relationship to albacore tuna gape limits.

```{r Albacore diet data + probable life stage estimation}

alb_global_prob = alb_global %>%
  inner_join(prey_length_ratiosp, by=c("prey_sp", "prey_family", "prey_order", "prey_class")) %>%
  #In using an inner_join I am selecting only rows for which we have all length-based data
  mutate(prey_length = if_else(maxtl_use > 0, maxtl_use/10, gape_lmax)) %>%
  mutate(prey_frac = 100*prey_length/l_adult) %>%
  mutate(prey_proba = if_else(prey_frac < larva_adult_use, "larva", if_else(prey_frac<juve_adult_use, "juvenile", "adult"))) %>%
  mutate(life_stage = if_else(prey_age_reported_1 == "none", prey_proba, prey_age_reported_1)) #985 obs, 53 variables

#And remove troublesome NA's in dataset
#Because we need to give these a life stage, I am assigning them as "adult" because of their l_max typically was < gape limit.
alb_global_prob$life_stage[is.na(alb_global_prob$life_stage)] <- "adult"

## SAVE DATA
write.csv(alb_global_prob, here("data/output_data/alb_global_diet_prob.csv")) 
#985 observations 53 variables
#This include data for which we are able to make an assessment of probable life stage --> therefore will exclude coarse taxonomic classifications

```

The following summarises the max observed FO, N and M for each species in the dataset AND for each life stage likely consumed.

```{r Alb probable diet metrics summary}

#All combinations of life stages and species consumed
alb_global_prob_metrics = alb_global_prob %>%
  dplyr::select(prey_class:prey_sp, life_stage, pFOuse, pN, pM) %>%
  group_by(prey_sp, life_stage) %>%
  dplyr::summarize(maxFO = max(pFOuse, na.rm=TRUE), maxN = max(pN, na.rm=TRUE), maxM = max(pM, na.rm=TRUE))

## SAVE DATA BELOW

```

AND we need to select the max observed.

```{r Albacore prey life stages lists}
#Need to summarise prey life stages and attach trait data
#unique(as.factor(alb_global_prob_metrics$prey_sp)) #there are 303 unique species
#So we should obtain 300 species in the following code.

#The most common life stage and species consumed!!
alb_global_prob_metrics = alb_global_prob_metrics %>%
  group_by(prey_sp) %>%
  top_n(1, maxFO) %>%
  sample_n(1)

write.csv(alb_global_prob_metrics, here("data/output_data/alb_global_prob_metrics"))

```

*Note: the maxFO, maxN or maxM are related to the study, they are typically the same between both sets of summary data for adult and probable life stage.

## DATA SETS NEEDED


## DATA JOINS

### Prey traits - adult life stage

```{r Prey adult traits + morph + qual data, include=FALSE, eval = FALSE}

#Mixed categorical, and other traits

prey_adults_merge = prey_trait_adult %>% 
  left_join(prey_morph_sum, by="prey_sp") %>% 
  left_join(prey_qual_sum, by="prey_sp") %>%
  left_join(alb_global_ad_metrics, by="prey_sp")

#308 spp and ~31 traits
glimpse(prey_adults_merge)

#Save
write.csv(prey_adults_merge, here("data/output_data/prey_adults_traits.csv"))

#Reload
#prey_adults_merge = read.csv(here("data/output_data/prey_adults_traits.csv"))

```

### Prey traits - probable life stage

```{r Prey probable list + traits + morph + qual data, include=FALSE, eval = FALSE}

prey_probable_merge = alb_global_prob_metrics %>%
  left_join(prey_trait_select, by=c("prey_sp", "life_stage")) %>% 
  left_join(prey_morph_sum, by="prey_sp") %>% 
  left_join(prey_qual_sum, by="prey_sp") %>%
  dplyr::select(prey_class:prey_family, prey_sp:life_stage, vert_habitat:percent_lipid, maxFO:maxM)

#303 spp and ~28 traits
glimpse(prey_probable_merge)

write.csv(prey_probable_merge, here("data/output_data/prey_probable_traits.csv"))

#prey_probable_merge = read.csv(here("data/prey_probable_traits.csv"))

```

#### Covariates for analyses

```{r Covariates}

alb_covars = alb_global %>%
  dplyr::select(StudyID, CiteYear, OceanBasin, OceanBasinQ, LocatName, LocatLatitude:pred_life_est) %>% #, stomachs_used
  distinct(StudyID, CiteYear, OceanBasin, OceanBasinQ, LocatName, LocatLatitude, LocatLongitude, YearEnd, MonthRange, stomachs_used, pred_life) %>% 
  #, stomachs_used
  mutate(lat_cat = if_else(abs(LocatLatitude) > 23.43662, "temperate", "tropical"),
         pred_life=case_when(pred_life == "juvenile" ~ "juvenile",
                             pred_life == "juvenile, adult" ~ "mixed",
                             pred_life == "adult" ~ "adult"))

write.csv(alb_covars, here("data/output_data/alb_covars.csv"))

```

Could edit the Longhurst doc to separate biomes and codes as per province definitions:
https://en.wikipedia.org/wiki/Longhurst_code#:~:text=Longhurst%20code%20refers%20to%20a,unique%20set%20of%20environmental%20conditions 

```{r Covars + Longhurst}
#Covars + Longhurst

alb_covars_geog <- read.csv(here("data/output_data/alb_covars_longhurst.csv")) %>%
  dplyr::select(StudyID, lat_cat:code) %>%
  group_by(StudyID) %>%
  sample_n(1)
names(alb_covars_geog)

#Need to merge this with diet data first overall data then probable prey consumed
unique(alb_covars_geog$code)

#check names match
unique(sort(alb_covars_geog$StudyID)) == unique(sort(alb_global$StudyID)) #they do
```


**NOTE** Below we include the adult cluster number but need to use the cluster number from probable life stage if we stick to this for final analysis.

#### Diet data - frequency of occurrence

```{r Covars + Longhurst + cluster + Diet data F}

#Covars + Longhurst + Diet data FO

alb_global_diet = alb_global %>%
  left_join(alb_covars_geog) #looks good #, by = "StudyID"

#Upload the cluster info: 
##This is where we choose to either upload cluster data from adult-based or probable-consumed cluster analysis.
alb_adult_traits = read.csv(here("data/output_data/adult.prey.clusternum.habgreg.aggl.k9.csv")) %>%
  dplyr::select(prey_sp:adult.clust.num)

#Extract %FO data --> filter for only species that we have adult traits + cluster info for --> transform to wide data for MV GLMs + trait GLMs
alb_global_dietfo = alb_global_diet %>%
  dplyr::select(StudyID:YearEnd, lat_cat:code, pred_life, pred_flmean, prey_sp, pFOuse) %>%
  filter(pFOuse >= 1) %>% #set this value to include or exclude species
  filter(prey_sp %in% alb_adult_traits$prey_sp) %>%
  group_by(prey_sp) %>%
  dplyr::mutate(grouped_id = row_number()) %>%
  spread(prey_sp, pFOuse, fill = 0, convert = FALSE)

#Before spreading data - range(alb_global_dietfo$pFOuse) 
#0 to 100 need to exclude zeros as they are not true zeros --> 0.1 to 100%

#Check covariate representation
#summary(as.factor(alb_global_dietfo$lat_cat))
#summary(as.factor(alb_global_dietfo$code))
#summary(as.factor(alb_global_dietfo$OceanBasin))
#summary(as.factor(alb_global_dietfo$OceanBasinQ))

write.csv(alb_global_dietfo, here("data/output_data/alb_global_wide_dietfo.csv"))

```

#### Diet data - abundance

```{r Covars + Longhurst + Diet data N}
#Extract for %N data
alb_global_dietn = alb_global_diet %>%
  dplyr::select(StudyID:YearEnd, lat_cat:code, pred_life, pred_flmean, prey_sp, pN) %>%
  filter(pN > 0) %>%
  filter(prey_sp %in% alb_adult_traits$prey_sp) %>%
  group_by(prey_sp) %>%
  dplyr::mutate(grouped_id = row_number()) %>%
  spread(prey_sp, pN, fill = 0, convert = FALSE)

#Check covariate representation
#summary(as.factor(alb_global_dietn$lat_cat))
#summary(as.factor(alb_global_dietn$code))
#summary(as.factor(alb_global_dietn$OceanBasin))
#summary(as.factor(alb_global_dietn$OceanBasinQ))

write.csv(alb_global_dietn, here("data/output_data/alb_global_wide_dietn.csv"))
```

#### Diet data - biomass

```{r Covars + Longhurst + Diet data M}
#Extract for %M data
#range(alb_global_dietm$pM) #0 to 95.2, need to exclude zeros as not true zeros

alb_global_dietm = alb_global_diet %>%
  dplyr::select(StudyID:YearEnd, lat_cat:code, pred_life, pred_flmean, prey_sp, pM) %>%
  filter(pM > 0) %>%
  filter(prey_sp %in% alb_adult_traits$prey_sp) %>%
  group_by(prey_sp) %>%
  dplyr::mutate(grouped_id = row_number()) %>%
  spread(prey_sp, pM, fill = 0, convert = FALSE)

#Check covariate representation
#summary(as.factor(alb_global_dietm$lat_cat))
#summary(as.factor(alb_global_dietm$code))
#summary(as.factor(alb_global_dietm$OceanBasin))
#summary(as.factor(alb_global_dietm$OceanBasinQ))

write.csv(alb_global_dietm, here("data/output_data/alb_global_wide_dietm.csv"))

```


